Large squeezing in multicomponent trapped Bose-Einstein condensates

### 2013-06-04 large_number_squeezing_with_atoms.tex
Fix figures, plus minor typo fixes

\begin{figure}
  \centering
   \includegraphics[width=7cm]{figures/volume_effect_on_squeezing.eps}
\caption{\textbf{This figure is not yet the 'winner winner chicken dinner' result. } Demonstration of high squeezing from a large volume, 3D trap.  \textbf{Simulation parameters.}}
 \label{figVolumeSqueezingEffects}
\end{figure}

@@

Beginning with the Hamiltonian $H_1$, during the period $(0, t_1)$ mode $\hat{a}$ will evolve as
\begin{equation}
\hat{a}_1 = e^{ i H_1 t/ \hbar} \, \hat{a}_0 \, e^{-i H_1 t/ \hbar}.
\end{equation}
Utilizing the Hadamard lemma
\begin{equation}
e^X Y e^{-X} = Y + [X,Y] + \frac{1}{2!}[X,[X,Y]] + \frac{1}{3!}[X,[X,[X,Y]]] + \ldots
\label{eqHadamard}
\end{equation}
we obtain 
\begin{eqnarray}
\hat{a}_1 &=& \hat{a}_0 - i \Omega t_1 \hat{b}_0 + \frac{(i \Omega t_1)^2}{2!} \hat{a}_0 - \frac{(i \Omega t_1)^3}{3!} \hat{b}_0 + \frac{(i \Omega t_1)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
          &=& \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0.
\label{eqa1}
\end{eqnarray}
Similarly, we find
\begin{equation}
\hat{b}_1 = \cos (\Omega t_1) \hat{b}_0 - i \sin (\Omega t_1) \hat{a}_0.
\label{eqb1}
\end{equation}
From the form of (\ref{eqa1}) and (\ref{eqb1}) it is clear that as both modes $\hat{a}$ and $\hat{b}$ began in coherent states, they will remain in coherent states, with only their amplitudes changing. Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$.

@@

As $\hat{a}^{\dagger}_1 \hat{a}_1$ and $\hat{b}^{\dagger}_1 \hat{b}_1$ commute with the Hamiltonian $H_2$, they are constants of motion so we have
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}_1 |\alpha, \beta \rangle = |\alpha|^2 = n_a
\end{equation}
%
\begin{equation}
\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = |\beta|^2 = n_b
\end{equation}
%
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}_1 \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = |\alpha|^2 |\beta|^2 = n_a n_b
\end{equation}
Furthermore, $\hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2$ and $\hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2$ also commute with $H_2$, giving
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \hat{a}^{\dagger}_1 \hat{a}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}^{\dagger}_1 \hat{a}_1 \hat{a}_1 + \hat{a}^{\dagger}_1 \hat{a}_1 |\alpha, \beta \rangle = n_a^2 + n_a
\end{equation}
%
\begin{equation}
\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{b}^{\dagger}_1 \hat{b}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{b}^{\dagger}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \hat{b}_1 + \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = n_b^2 + n_b
\end{equation}
The remaining terms are not constants of motion, so we proceed by making use of (\ref{eqa2}) and (\ref{eqb2}). For $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$ we have
\begin{equation}
\hat{a}^{\dagger}_2 \hat{b}_2 = \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1. 
\end{equation}
To calculate the expectation value of this expression, the creation and annihilation operators must be normally ordered. Making use of the identity \cite{louisell}
\begin{equation}
e^{x \hat{a}^{\dagger} \hat{a}} = \sum_{n=0}^{\infty} \frac{(e^x - 1)^n} {n!} (\hat{a}^{\dagger})^n (\hat{a})^n
\end{equation}
allows us to compute expectation values of coherent states of the form
\begin{eqnarray}
\langle \alpha | e^{x \hat{a}^{\dagger} \hat{a}} |\alpha \rangle &=& \sum_{n=0}^{\infty} \frac{(e^x - 1)^n} {n!} |\alpha|^{2n} \nonumber \\
%
&=& \exp[|\alpha|^2(e^x -1)]. \label{eqExpectationValueOfExponential}
\end{eqnarray}
This gives
\begin{eqnarray}
\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle &=& \langle \alpha, \beta | \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1 | \alpha, \beta \rangle \nonumber \\
%
&=& \alpha^* \beta \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2 (e^{-i (\lambda_{bb} - \lambda_{ab})} -1)].
\end{eqnarray}

@@

When substituted into (\ref{eqNavariance}) and (\ref{eqNbvariance}) these expecation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}. To calculate the number difference variance, we note

>>>>

@@

Beginning with the Hamiltonian $H_1$, during the period $(0, t_1)$ mode $\hat{a}$ will evolve as
\begin{equation}
\hat{a}_1 = e^{ i \hat{H}_1 t/ \hbar} \, \hat{a}_0 \, e^{-i \hat{H}_1 t/ \hbar}.
\end{equation}
Utilizing the Hadamard lemma
\begin{equation}
e^{\hat{X}} \hat{Y} e^{-\hat{X}} = \hat{Y} + [\hat{X},\hat{Y}] + \frac{1}{2!}[\hat{X},[\hat{X},\hat{Y}]] + \frac{1}{3!}[\hat{X},[\hat{X},[\hat{X},\hat{Y}]]] + \ldots
\label{eqHadamard}
\end{equation}
we obtain 
\begin{equation}
\hat{a}_1 =  \cos (\Omega t_1) \hat{a}_0 -i \sin (\Omega t_1) \hat{b}_0.
\label{eqa1}
\end{equation}
Similarly, we find
\begin{equation}
\hat{b}_1 = \cos (\Omega t_1) \hat{b}_0 - i \sin (\Omega t_1) \hat{a}_0.
\label{eqb1}
\end{equation}
From the form of (\ref{eqa1}) and (\ref{eqb1}) it is clear that as both modes $\hat{a}$ and $\hat{b}$ began in coherent states, they will remain in coherent states, with only their amplitudes changing. Specifically, at time $t_1$ the system is in a product of coherent states, which we denote $|\alpha, \beta\rangle$ with $\alpha=\sqrt{n_a}$ and $\beta=-i\sqrt{n_b}$.

@@

As $\hat{a}^{\dagger}_1 \hat{a}_1$ and $\hat{b}^{\dagger}_1 \hat{b}_1$ commute with the Hamiltonian $\hat{H}_2$, they are constants of motion so we have
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}_1 |\alpha, \beta \rangle = |\alpha|^2 = n_a
\end{equation}
%
\begin{equation}
\langle \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = |\beta|^2 = n_b
\end{equation}
%
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}_1 \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = |\alpha|^2 |\beta|^2 = n_a n_b
\end{equation}
Furthermore, $\hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2$ and $\hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2$ also commute with $H_2$, giving
\begin{equation}
\langle \hat{a}^{\dagger}_2 \hat{a}_2 \hat{a}^{\dagger}_2 \hat{a}_2 \rangle = \langle \hat{a}^{\dagger}_1 \hat{a}_1 \hat{a}^{\dagger}_1 \hat{a}_1 \rangle = \langle \alpha, \beta | \hat{a}^{\dagger}_1 \hat{a}^{\dagger}_1 \hat{a}_1 \hat{a}_1 + \hat{a}^{\dagger}_1 \hat{a}_1 |\alpha, \beta \rangle = n_a^2 + n_a
\end{equation}
%
\begin{equation}
\langle \hat{b}^{\dagger}_2 \hat{b}_2 \hat{b}^{\dagger}_2 \hat{b}_2 \rangle = \langle \hat{b}^{\dagger}_1 \hat{b}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \rangle = \langle \alpha, \beta | \hat{b}^{\dagger}_1 \hat{b}^{\dagger}_1 \hat{b}_1 \hat{b}_1 + \hat{b}^{\dagger}_1 \hat{b}_1 |\alpha, \beta \rangle = n_b^2 + n_b
\end{equation}
The remaining terms are not constants of motion, so we proceed by making use of (\ref{eqa2}) and (\ref{eqb2}). For $\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle$ we have
\begin{equation}
\hat{a}^{\dagger}_2 \hat{b}_2 = \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1. 
\end{equation}
To calculate the expectation value of this expression, the creation and annihilation operators must be normally ordered. Making use of the identity \cite{louisell}
\begin{equation}
e^{x \hat{a}^{\dagger} \hat{a}} = \sum_{n=0}^{\infty} \frac{(e^x - 1)^n} {n!} (\hat{a}^{\dagger})^n (\hat{a})^n
\end{equation}
allows us to compute expectation values of coherent states of the form
\begin{eqnarray}
\langle \alpha | e^{x \hat{a}^{\dagger} \hat{a}} |\alpha \rangle &=& \sum_{n=0}^{\infty} \frac{(e^x - 1)^n} {n!} |\alpha|^{2n} \nonumber \\
%
&=& \exp[|\alpha|^2(e^x -1)]. \label{eqExpectationValueOfExponential}
\end{eqnarray}
This gives
\begin{eqnarray}
\langle \hat{a}^{\dagger}_2 \hat{b}_2 \rangle &=& \langle \alpha, \beta | \hat{a}^{\dagger}_1 e^{-i (\lambda_{bb} - \lambda_{ab}) \hat{b}^{\dagger}_1 \hat{b}_1} e^{i (\lambda_{aa} - \lambda_{ab}) \hat{a}^{\dagger}_1 \hat{a}_1} \hat{b}_1 | \alpha, \beta \rangle \nonumber \\
%
&=& \alpha^* \beta \exp[|\alpha|^2(e^{i (\lambda_{aa} - \lambda_{ab})} -1)] \exp[|\beta|^2 (e^{-i (\lambda_{bb} - \lambda_{ab})} -1)].
\end{eqnarray}

@@

When substituted into (\ref{eqNavariance}) and (\ref{eqNbvariance}) these expectation values give the absolute number variances stated in Section \ref{secTwoModeAnalytic}. To calculate the number difference variance, we note
